---
title: "Probabilistic Reconciliation via Conditioning with `bayesRecon`"
author: "Nicolò Rubattu, Giorgio Corani, Dario Azzimonti, Lorenzo Zambon"
date: "2023-07-26"
lang: "en"
output: html_vignette
bibliography: references.bib
cite:
- '@zambon2022efficient'
- '@zambon2023properties'
- '@corani2023probabilistic'
- '@corani2021probabilistic'
vignette: >
  %\VignetteIndexEntry{Probabilistic Reconciliation via Conditioning with `bayesRecon`}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval=TRUE ### !!!! set to FALSE here to render only the text !!!!
)
set.seed(42)
```

```{r klippy, echo=FALSE, include=TRUE, eval=FALSE}
klippy::klippy(position = c('top', 'right'), tooltip_message = 'Copy', tooltip_success = 'Done', color="black")
```

# Introduction

This vignette shows how to use the `bayesRecon` package for the *probabilistic reconciliation* of hierarchical forecasts. We provide three examples:

1. **Carpart**: we build a temporal hierarchy on a count time series and we reconcile it via the Bottom-Up Importance Sampling (BUIS)  algorithm.

2. **M3**: we build a temporal hierarchy on a smooth time series and we reconcile it using Gaussian reconciliation.

3. **Infants mortality**: we reconcile cross-sectional forecasts.

The implemented reconciliation algorithms are discussed in:

* Zambon, L., Azzimonti, D., & Corani, G. (2022). Efficient probabilistic reconciliation of forecasts for real-valued and count time series.

* Zambon, L., Agosto, A., Giudici, P., & Corani, G. (2023). Properties of the reconciled distributions for Gaussian and count forecasts.

* Corani, G., Azzimonti, D., & Rubattu, N. (2023). Probabilistic reconciliation of count time series.

* Corani, G., Azzimonti, D., Augusto, J. P., & Zaffalon, M. (2021). Probabilistic reconciliation of hierarchical forecast via Bayes' rule.

# Setup

You can install the stable version of `bayesRecon` on [R CRAN](https://cran.r-project.org/package=bayesRecon).

```{r install, eval=FALSE}
install.packages('bayesRecon', dependencies = TRUE)
```

Load the package.
```{r load}
library(bayesRecon)
```

# Example 1: Carpart
In this example, we consider a *monthly* time series of car part sales, from Jan. 1998 to Mar. 2002. \
This is actually the time series #2655 of the *carparts* dataset [@hyndman2008forecasting]. It is available from `bayesRecon::carpart`.

The time series consists of **count** data, and the distribution of its values is skewed. Hence, we  produce base forecasts using a method specialized for  count time series.

```{r carpart-plot, dpi=300, out.width = "100%", fig.align='center', fig.cap="**Figure 1**: Carpart - monthly car part sales.", fig.dim = c(6, 3)}
layout(mat = matrix(c(1, 2), nrow = 1, ncol = 2), widths = c(2, 1))
plot(carpart, xlab = "Time", ylab = "Car part sales", main = NULL)
hist(carpart, xlab = "Car part sales", main = NULL)
```
<br><br>
We divide the time series into train and test;  we want to forecast the test set, i.e., the last 12 observations (1 year ahead).
```{r train-test}
train <- window(carpart, end = c(2001, 3))
test <- window(carpart, start = c(2001, 4))
```

<!-- In order to do hierarchical forecasting: -->

<!-- 1.    we build a temporal hierarchy; -->
<!-- 2.    we compute the *base forecasts*) at each temporal scale. -->

We now build the temporal hierarchy and compute the *base forecasts*.

We build the hierarchy using the `temporal aggregation` function. We aggregate at *2-Monthly*, *Quarterly*, *4-Monthly*, *Biannual*, and *Annual* by specifying the `agg_levels` argument.

``` {r temp-agg}
train.agg <- bayesRecon::temporal_aggregation(train, agg_levels = c(2, 3, 4, 6, 12))
levels <- c("Annual", "Biannual", "4-Monthly", "Quarterly", "2-Monthly", "Monthly")
names(train.agg) <- levels
```

The function returns a list of aggregated time series, ordered from the top (most aggregated) to the bottom (most disagreggated).
``` {r temp-agg-plot, dpi=300, fig.show="hold", out.width="100%", out.heigth="100%", fig.align='center', fig.cap="**Figure 2**: Carpart - visualization of the aggregated time series.", fig.dim=c(6,3.5)}
par(mfrow = c(2, 3), mai = c(0.6, 0.6, 0.5, 0.5))
for (l in levels) {
  plot(train.agg[[l]], xlab = "Time", ylab = "Car part sales", main = l)
}
```
<br><br>
We use the  `glarma` package ([R CRAN](https://cran.r-project.org/package=glarma)) to generate the  forecasts. This package is specialized in forecasting count time series and it is based on
a Generalized Linear Autoregressive Moving Average model.
In this example, we adopt a Poisson predictive distribution.

The models produce forecasts for 1 year ahead;
<!-- *h* steps ahead according to the temporal level. -->
this is *h=12* steps ahead for the monthly level,  *h=4* steps ahead  for the quarterly level, etc.
Each forecast is constituted by $D$ integer samples;
we set $D=20000$.

Below we use a for loop to iterate over the levels of the temporal hierarchy.
For each level, we fit an independent model; then  we obtain 
the  samples of the forecasts using the function `glarma::forecast.glarma`.
Overall we have 28 predictive distributions (one for *Annual*, two for *Biannual*, etc.).
We collect the samples of the 28 distributions in a list.

The following lines of code take ~ 30 seconds on a personal computer.

``` {r hier-fore, cache=TRUE}
# install.packages("glarma", dependencies = TRUE)
#library(glarma)

fc.samples <- list()
D <- 20000
fc.count <- 1
# iterating over the temporal aggregation levels
for (l in seq_along(train.agg)) {
  f.level <- frequency(train.agg[[l]])
  print(paste("Forecasting at ", levels[l], "...", sep = ""))
  # fit an independent model for each aggregation level
  model <- glarma::glarma(
    train.agg[[l]],
    phiLags = if (f.level == 1) 1 else 1:(min(6, f.level - 1)),
    thetaLags = if (f.level == 1) NULL else f.level,
    X = cbind(intercept = rep(1, length(train.agg[[l]]))),
    offset = cbind(intercept = rep(0, length(train.agg[[l]]))),
    type = "Poi"
  )
  # forecast 1 year ahead
  h <- f.level
  tmp <- matrix(data = NA, nrow = h, ncol = D)
  for (s in (1:D)) {
    # each call to 'forecast.glarma' returns a simulation path
    tmp[, s] <- glarma::forecast(
      model,
      n.ahead = h,
      newdata = cbind(intercept = rep(1, h)),
      newoffset = rep(0, h)
    )$Y
  }
  # collect the forecasted samples
  for (i in 1:h) {
    fc.samples[[fc.count]] <- tmp[i, ]
    fc.count <- fc.count + 1
  }
}

```


Base forecasts do not sum up correctly. In order to address incoherence we can now use a reconciliation algorithm.

First, we obtain the summing matrix $\mathbf{S}$ using the function `get_reconc_matrices`. 
It requires: 

*   the aggregation factors used before to build the hierarchy, $\{2, 3, 4, 6, 12\}$ in this example;
*   the length of the forecasting horizon for the bottom time series (*h=12* in this example).

``` {r S}
recon.matrices <- bayesRecon::get_reconc_matrices(agg_levels = c(2, 3, 4, 6, 12), h = 12)
# Summing matrix
S <- recon.matrices$S 
# A matrix
A <- recon.matrices$A 
```

Then, we use the function `reconc_BUIS` which implements the Bottom-Up Important Sampling (BUIS) algorithm.
It requires the $\mathbf{S}$ matrix, the *base forecasts*, the  type of the base forecasts ("samples" in this example) and the specification of the base forecasts distributions ("discrete" in this example, since the predictive distributions are Poisson).

The `in_type` argument is a string with two possible values: "samples" if the base forecasts are in the form of samples, "params" if the base forecasts are in the form of estimated parameters. The `distr` argument is a string describing the type of base forecasts. It can be: "continuous" or "discrete" if `in_type = "samples"`; "gaussian", "poisson" or "nbinom" if `in_type = "params"`. 

``` {r reconc}
recon.res <- bayesRecon::reconc_BUIS(
  S,
  base_forecasts = fc.samples,
  in_type = "samples",
  distr = "discrete",
  seed = 42
)
```
The function returns samples from the  *reconciled forecast distribution*.
```{r res}
reconciled_samples <- recon.res$reconciled_samples
dim(reconciled_samples)

```

We now compute the Mean Absolute Error (MAE) and the Continuous Ranked Probability Score (CRPS) for the bottom (i.e., *monthly*) time series.  For CRPS, we use `crps_sample` from the package `scoringRules` ([R CRAN](https://cran.r-project.org/package=scoringRules)).

```{r metrics}
# install.packages("scoringRules", dependencies = TRUE)
library(scoringRules)

ae.fc <- list()
ae.reconc <- list()
crps.fc <- list()
crps.reconc <- list()
for (h in 1:length(test)) {
  y.hat_ <- median(fc.samples[[nrow(A) + h]])
  y.reconc_ <- median(recon.res$bottom_reconciled_samples[, h])
  # Compute Absolute Errors
  ae.fc[[h]] <- abs(test[h] - y.hat_)
  ae.reconc[[h]] <- abs(test[h] - y.reconc_)
  # Compute Continuous Ranked Probability Score (CRPS)
  crps.fc[[h]] <-
    scoringRules::crps_sample(y = test[h], dat = fc.samples[[nrow(A) + h]])
  crps.reconc[[h]] <-
    scoringRules::crps_sample(y = test[h], dat = recon.res$bottom_reconciled_samples[, h])
}

mae.fc <- mean(unlist(ae.fc))
mae.reconc <- mean(unlist(ae.reconc))
crps.fc <- mean(unlist(crps.fc))
crps.reconc <- mean(unlist(crps.reconc))
metrics <- data.frame(
  row.names = c("MAE", "CRPS"),
  base.forecasts = c(mae.fc, crps.fc),
  reconciled.forecasts = c(mae.reconc, crps.reconc)
)
metrics

```

Note the improvements with respect to the base forecast on both MAE and CRPS metrics.

# Example 2: M3

We now consider a *monthly* time series (N1485) from the M3 forecasting competition [@makridakis2000m3].
It is available from `bayesRecon::M3sample`.

```{r m3-plot, dpi=300, out.width = "100%", fig.align='center', fig.cap="**Figure 3**: M3 - N1485 time series.", fig.dim = c(6, 3)}
plot(M3example$train, xlab = "Time", ylab = "y", main = "N1485")
```
<br>
We build the temporal hierarchy using the `temporal_aggregation` function. We aggregate at *2-Monthly*, *Quarterly*, *4-Monthly*, *Biannual*, and *Annual*. As we do not specify the `agg_levels` argument, the function generates
all the factors of the time series frequency  (i.e. `frequency(M3example$train)` ).

```{r m3-agg} 
train.agg <- bayesRecon::temporal_aggregation(M3example$train)
levels <- c("Annual", "Biannual", "4-Monthly", "Quarterly", "2-Monthly", "Monthly")
names(train.agg) <- levels
```


We use  the ets models from  `forecast` ([R CRAN](https://cran.r-project.org/package=forecast)) to generate forecast  for each temporal level, for each *h*. 
The predictive distribution is  Gaussian and we predict
the next 18 months (the value of *h* is different for each temporal level). 

```{r m3-fore}
# install.packages("forecast", dependencies = TRUE)
library(forecast)

H <- length(M3example$test)
H

fc <- list()
level.idx <- 1
fc.idx <- 1
for (level in train.agg) {
  level.name <- names(train.agg)[level.idx]
  # fit an ETS model for each temporal level
  model <- ets(level)
  # generate forecasts for each level within 18 months
  h <- floor(H / (12 / frequency(level)))
  print(paste("Forecasting at ", level.name, ", h=", h, "...", sep = ""))
  level.fc <- forecast(model, h = h)
  # save mean and sd of the gaussian predictive distribution
  for (i in 1:h) {
    fc[[fc.idx]] <- c(level.fc$mean[[i]],
                      (level.fc$upper[, "95%"][[i]] - level.fc$mean[[i]]) / qnorm(0.975))
    fc.idx <- fc.idx + 1
  }
  level.idx <- level.idx + 1
}
```


Before reconciling the forecasts,  we obtain the matrix $\mathbf{S}$ using the function `get_reconc_matrices`.

```{r m3-rmat, dpi=300, out.width = '70%', fig.align='center', fig.cap="**Figure 4**: M3 - The aggregating matrix A (red=1, yellow=0).", fig.dim = c(8, 8)}
rmat <- get_reconc_matrices(agg_levels = c(2, 3, 4, 6, 12), h = 18)

par(mai = c(1,1,0.5,0.5))
image(1:ncol(rmat$A), 1:nrow(rmat$A), 
      t(apply(t(rmat$A),1,rev)), 
      xaxt='n', yaxt='n', ylab = "", xlab = levels[6])
axis(1, at=1:ncol(rmat$A), label=1:ncol(rmat$A), las=1)
axis(2, at=c(23,22,19,15,9), label=levels[1:5], las=2)
```
<br>
We now use the function `reconc_gaussian` which implements the closed-form Gaussian reconciliation,
and the function `reconc_BUIS` which implements the Bottom-Up Important Sampling (BUIS) algorithm. 
We can check that the algorithms return consistent results, apart from minor numerical differences due to the sampling randomness of BUIS, which get smaller as we increase the `num_samples` argument of `reconc_BUIS`.

```{r m3-reco} 
recon.gauss <- bayesRecon::reconc_gaussian(
  S = rmat$S,
  base_forecasts.mu = sapply(fc, "[[", 1),
  base_forecasts.Sigma = diag(sapply(fc, "[[", 2)) ^ 2
)

reconc.buis <- bayesRecon::reconc_BUIS(
  S = rmat$S,
  base_forecasts = fc,
  in_type = "params",
  distr = "gaussian",
  num_samples = 20000,
  seed = 42
)

# check that the algorithms return consistent results
round(rbind(
  c(recon.gauss$upper_reconciled_mean, recon.gauss$bottom_reconciled_mean),
  rowMeans(reconc.buis$reconciled_samples)
))
```

We now show the *base forecasts* and the *reconciled forecasts*.

```{r m3-plotfore, dpi=300, out.width = "100%", fig.align='center', fig.cap="**Figure 5**: M3 - visualization of base and reconciled forecasts. The black line is the actual data (dashed is the test). The orange line is the forecasted mean, the blu line the reconciled mean. Shadow regions show the 95% prediction intervals.", fig.dim = c(6, 4)}
yhat.mu <- tail(sapply(fc, "[[", 1), 18)
yhat.sigma <- tail(sapply(fc, "[[", 2), 18)
yhat.hi95 <- qnorm(0.975, mean = yhat.mu, sd = yhat.sigma)
yhat.lo95 <- qnorm(0.025, mean = yhat.mu, sd = yhat.sigma)
yreconc.mu <- rowMeans(reconc.buis$bottom_reconciled_samples)
yreconc.hi95 <- apply(reconc.buis$bottom_reconciled_samples, 1, 
                      function(x) quantile(x, 0.975))
yreconc.lo95 <- apply(reconc.buis$bottom_reconciled_samples, 1, 
                      function(x) quantile(x, 0.025))

ylim_ <- c(min(M3example$train, M3example$test, yhat.lo95, yreconc.lo95) - 1, 
           max(M3example$train, M3example$test, yhat.hi95, yreconc.hi95) + 1)

plot(M3example$train, xlim = c(1990, 1995.6), ylim = ylim_, 
     ylab = "y", main = "N1485 Forecasts")
lines(M3example$test, lty = "dashed")
lines(ts(yhat.mu, start = start(M3example$test), frequency = 12), 
      col = "coral", lwd = 2)
lines(ts(yreconc.mu, start = start(M3example$test), frequency = 12), 
      col = "blue2", lwd = 2)
xtest <- time(M3example$test)
polygon(c(xtest, rev(xtest)), c(yhat.mu, rev(yhat.hi95)), 
        col = "#FF7F5066", border = "#FF7F5066")
polygon(c(xtest, rev(xtest)), c(yhat.mu, rev(yhat.lo95)), 
        col = "#FF7F5066", border = "#FF7F5066")
polygon(c(xtest, rev(xtest)), c(yreconc.mu, rev(yreconc.hi95)), 
        col = "#0000EE4D", border = "#0000EE4D")
polygon(c(xtest, rev(xtest)), c(yreconc.mu, rev(yreconc.lo95)), 
        col = "#0000EE4D", border = "#0000EE4D")
```

# Example 3: Infants mortality

In this example, we consider the *infantgts* hierarchical time series [@hyndman2011optimal].
It is available via `bayesRecon::infantMortality`.

It is a *yearly* grouped **cross-sectional** time series dataset, from 1901 to 2003, of infant mortality counts (deaths) in Australia, disaggregated by state and sex (male and female). 
State codes refer to: New South Wales (NSW), Victoria (VIC), Queensland (QLD), South Australia (SA), Western Australia (WA), Northern Territory (NT), Australian Capital Territory (ACT), and Tasmania (TAS).

We want to forecast the next year, 2004. We use the auto.arima forecasting models from the package `forecast` ([R CRAN](https://cran.r-project.org/package=forecast)) to generate Gaussian predictions for each node of the hierarchy. 

For each model, we collect the residuals to take into account correlations. 

```{r infants-forecasts}
# install.packages("forecast", dependencies = TRUE)
library(forecast)

fc <- list()
residuals <- matrix(NA,
                    nrow = length(infantMortality$total),
                    ncol = length(infantMortality))
fc.idx <- 1
for (s in infantMortality) {
  s.name <- names(infantMortality)[fc.idx]
  print(paste("Forecasting at ", s.name, "...", sep = ""))
  # fit an auto.arima model and forecast with h=1
  model <- auto.arima(s)
  s.fc <- forecast(model, h = 1)
  # save mean and sd of the gaussian predictive distribution
  fc[[s.name]] <- c(s.fc$mean,
                    (s.fc$upper[, "95%"][[1]] - s.fc$mean) / qnorm(0.975))
  residuals[, fc.idx] <- s.fc$residuals
  fc.idx <- fc.idx + 1
}
```

Build the $\mathbf{S}$ matrix.

```{r infants-s, dpi=300, out.width = '70%', fig.align='center', fig.cap="**Figure 6**: Infants mortality - The aggregating matrix A (red=1, yellow=0).", fig.dim = c(8, 8)}
# we have 16 bottom time series, and 11 upper time series
A <- matrix(data = c(1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,
                     1,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,
                     0,0,1,1,0,0,0,0,0,0,0,0,0,0,0,0,
                     0,0,0,0,1,1,0,0,0,0,0,0,0,0,0,0,
                     0,0,0,0,0,0,1,1,0,0,0,0,0,0,0,0,
                     0,0,0,0,0,0,0,0,1,1,0,0,0,0,0,0,
                     0,0,0,0,0,0,0,0,0,0,1,1,0,0,0,0,
                     0,0,0,0,0,0,0,0,0,0,0,0,1,1,0,0,
                     0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,1,
                     1,0,1,0,1,0,1,0,1,0,1,0,1,0,1,0,
                     0,1,0,1,0,1,0,1,0,1,0,1,0,1,0,1), byrow=TRUE, ncol = 16)
S <- rbind(A, diag(16))

# plot of A
par(mai = c(1.5,1,0.5,0.5))
image(1:ncol(A), 1:nrow(A), 
      t(apply(t(A),1,rev)), 
      xaxt='n', yaxt='n', ann=FALSE)
axis(1, at=1:ncol(A), label=names(infantMortality)[12:27], las=2)
axis(2, at=c(1:11), label=rev(names(infantMortality)[1:11]), las=2)
```
<br>

We can now use Gaussian reconciliation, to ensure *coherent* forecasts. 

We require an estimate of the covariance of the base forecasts. We can estimate it 
with the sample covariance of the in-sample errors. However, this is often 
a poor estimate due to the high variance, especially when the number of series in the hierarchy
is larger than the length of the series. @wickramasuriya2019optimal proposed an estimator 
of the covariance that shrinks the off-diagonal elements towards 0.

The function `bayesRecon::schaferStrimmer_cov` implements this estimator by following 
@schafer2005shrinkage to compute the optimal
shrinkage intensity.

```{r infants reconc}
# means
mu <- sapply(fc, "[[", 1)
# Shrinkage covariance
shrink.res <- bayesRecon::schaferStrimmer_cov(residuals)
print(paste("The estimated shrinkage intensity is", round(shrink.res$lambda_star, 3)))
Sigma <- shrink.res$shrink_cov
```

The function `reconc_gaussian` implements the closed-form Gaussian reconciliation. It requires the covariance matrix and the vector of means.

```{r infants-recon}
recon.gauss <- bayesRecon::reconc_gaussian(S,
                                           base_forecasts.mu = mu,
                                           base_forecasts.Sigma = Sigma)

# check coherence
(A %*% recon.gauss$bottom_reconciled_mean) - recon.gauss$upper_reconciled_mean
```

# References
<div id="refs"></div>
