---
title: "t-Reconciliation: Reconciliation of Gaussian forecasts by modelling the uncertainty on the covariance matrix"
author: "Dario Azzimonti, Chiara Carrara, Lorenzo Zambon, Giorgio Corani"
date: ""
lang: "en"
output: rmarkdown::html_vignette
bibliography: references.bib
cite:
- '@zambon2024properties'
- '@carrara2025'
vignette: >
  %\VignetteIndexEntry{t-Reconciliation: Reconciliation of Gaussian forecasts by modelling the uncertainty on the covariance matrix}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

This vignette shows how to perform *t-reconciliation*, a Bayesian approach to forecast reconciliation that explicitly accounts for uncertainty in the covariance matrix of residuals.

By exploiting the flexibility of Bayesian models to incorporate parameter uncertainty, we place an Inverse-Wishart prior on the covariance matrix. This choice allows the reconciliation to be derived in closed form, resulting in a reconciled predictive distribution that follows a multivariate t.


```{r load, message=FALSE, warning=FALSE}
# load the packages
library(bayesRecon)
library(forecast) # for base forecasts
library(ggplot2)  # for plot
```

# Loading the dataset

The example presented uses monthly Swiss overnight tourist stays data divided by Canton and aggregated over the whole country. This generates a cross-sectional hierarchical structure, consisting of two levels: the national total and the division by canton. The data is available in the package `bayesRecon` and can be loaded as `bayesRecon::swiss_tourism`. The raw data is publicly available on the official website of the Swiss Confederation at this [link](https://www.bfs.admin.ch/asset/en/px-x-1003020000_102).

The dataset spans the period from January 2005 to January 2025, comprising 241 monthly observations.

The time series exhibit strong seasonality. As an example, we plot the data for the top-level (aggregate) series in **Figure 1**.

```{r swiss-tourism-plot, dpi=300, out.width = "100%", fig.align='center', fig.cap="**Figure 1**: Swiss tourism - monthly Swiss overnight stays.", fig.dim = c(8, 4)}
# save all time series 
Y = swiss_tourism$ts

autoplot(Y[,1], ylab = "Overnight stays in Switzerland",linewidth=1)+
  scale_y_continuous(labels = function(x) paste0(formatC(x / 1e6, format = "g"), "M"))
```

We load the aggregation matrix and select the length of the training set. The length of the training set can be selected within the range [26, 240] observations.

```{r Swiss tourism}
A = swiss_tourism$agg_mat

n_b = ncol(A)
n_u = nrow(A)
n = n_b + n_u

# Frequency is monthly:
print(frequency(Y))

# Select the length of the training set and the forecast horizon
L = 60    
h = 1    

train = window(Y, end = time(Y)[L])
actuals = window(Y, start= time(Y)[L + 1], end = time(Y)[L + h])
```

# Forecasts

The base forecasts are generated using the `ets()` function from the `forecast` package, which fits individual exponential smoothing models to each time series.

For each time series forecast, we save the base forecast mean (`base_fc`) and the residuals from the fitted model (`res`), which are used to estimate the covariance matrix of the forecast errors.


```{r Base-forecasts and residuals computation}
# Compute base forecasts and residuals for each time series
base_fc = rep(1, n)
res = matrix(NA, ncol = n, nrow = L)
for (i in 1:n){
    models = forecast::ets(train[,i], model = "AZZ")
    f = forecast(models, h = h)
    base_fc[i] = f$mean
    res[, i] = models$residuals
}
```

# Reconciliation 

We can now reconcile the forecasts with the function `reconc_t()`, which implements the t-Reconciliation method. 

This function takes as input the aggregation matrix `A`, the base forecast means `base_fc_mean`, the training data `y_train` and the model residuals `residuals` and returns the parameters of the reconciled forecasts, which are distributed as a multivariate-t.

```{r t-reconciliation}
t_rec_results = reconc_t(A, base_fc_mean = base_fc, 
                         y_train = train, 
                         residuals = res,
                         return_parameters = TRUE)
```

For the selected data window, we save the base forecasts (Base) and we further compute the reconciled forecasts with two additional methods: Bottom-Up reconciliation (BU), standard Gaussian reconciliation (Gauss/MinT).

```{r Base}
#Base forecasts
Base_reconciled_mean = base_fc
Sigma = crossprod(res)/nrow(res)
# we apply a small amount of shrinkage to the covariance of the residuals for stability
Base_cov_mat = (1 - 10^-4)* Sigma + 10^-4 * diag(diag(Sigma))
```


```{r Bottom-up}
# Bottom-Up reconciliation
S = rbind(A, diag(rep(1, n_b)))  # summing matrix
BU_reconciled_mean = S %*% base_fc[(n_u + 1):n]
BU_cov_mat = S %*% Base_cov_mat[(n_u + 1):n, (n_u + 1):n] %*% t(S)
```

```{r Gaussian}
# Gaussian/MinT: compute reconciliation with bayesRecon
recon_gauss = reconc_gaussian(A, base_fc, residuals = res)
# save reconciled mean and covariance for the whole hierarchy
MinT_reconciled_mean = S %*% recon_gauss$bottom_rec_mean
MinT_cov_mat = S %*% recon_gauss$bottom_rec_cov %*% t(S)
```

# Comparison of results

For the selected data window, we compute and compare the predictive densities of the upper-level time series using four different methods: the base forecasts (Base), Bottom-Up reconciliation (BU), Minimum Trace reconciliation (MinT), and t-Reconciliation (t-Rec).

```{r compute uppers, echo=FALSE, eval=TRUE}

t_Rec_reconciled_mean <- S %*% t_rec_results$bottom_rec_mean
t_Rec_scale_par <- S %*% t_rec_results$bottom_rec_scale_matrix %*% t(S)
    
# Index of the upper variable to plot
i_upper <- 1  # change this if a different index is needed

# Extract distribution's parameters for each method
# BU
mu_BU <- as.numeric(BU_reconciled_mean[i_upper])
sd_BU <- sqrt(BU_cov_mat[i_upper, i_upper])

# MinT
mu_MinT <- as.numeric(MinT_reconciled_mean[i_upper])
sd_MinT <- sqrt(MinT_cov_mat[i_upper, i_upper])

# t-Rec
mu_tRec <- as.numeric(t_Rec_reconciled_mean[i_upper])
scale_tRec <- sqrt(t_Rec_scale_par[i_upper, i_upper])
df_tRec <- t_rec_results$bottom_rec_df

#Base
mu_Base <- as.numeric(Base_reconciled_mean[i_upper])
sd_Base <- sqrt(Base_cov_mat[i_upper, i_upper])

# Create a grid of x values
x_vals <- seq(min(mu_BU, mu_MinT, mu_tRec, mu_Base) - 4*max(sd_BU, sd_MinT, scale_tRec, sd_Base),
              max(mu_BU, mu_MinT, mu_tRec, mu_Base) + 4*max(sd_BU, sd_MinT, scale_tRec, sd_Base),
              length.out = 1000)

# Compute densities
dens_BU <- dnorm(x_vals, mean = mu_BU, sd = sd_BU)
dens_MinT <- dnorm(x_vals, mean = mu_MinT, sd = sd_MinT)
dens_tRec <- dt((x_vals - mu_tRec) / scale_tRec, df = df_tRec) /scale_tRec
dens_Base <- dnorm(x_vals, mean = mu_Base, sd = sd_Base)

max_y <- max(dens_BU, dens_MinT, dens_tRec, dens_Base)

```

```{r comparison-plot, echo=FALSE, eval=TRUE, dpi=300, out.width = "100%", fig.align='center', fig.cap="**Figure 2**: Predictive densities of the upper time series obtained with BU (yellow), MinT (purple), t-Rec (green) and Base (blue). The black triangle indicates the actual value.", fig.dim = c(8, 4), warning=FALSE}

# dens_df <- data.frame(
#   x = x_vals,
#   BU = dens_BU,
#   MinT = dens_MinT,
#   "t-Rec" = dens_tRec,
#   Base = dens_Base,
#   check.names = "FALSE"
# ) %>%
#   pivot_longer(cols = -x, names_to = "Method", values_to = "Density")
dens_df <- data.frame(
  x       = rep(x_vals, 4),
  Method  = rep(c("BU", "MinT", "t-Rec", "Base"), each = length(x_vals)),
  Density = c(dens_BU, dens_MinT, dens_tRec, dens_Base)
)

method_colors <- c(
  "BU" = "#FDE725",
  "MinT" = "#440154",
  "t-Rec" = "#29AF7F",
  "Base" = "#3B528B"
)

ggplot(dens_df, aes(x = x, y = Density, color = Method)) +
  geom_area(
    data = dens_df[dens_df$Method %in% c("MinT", "t-Rec"),],
    aes(fill = Method),
    position = "identity",
    alpha = 0.3,
    color = NA
  ) +
  geom_line(linewidth = 1.5) +
  scale_color_manual(values = method_colors) +
  scale_fill_manual(values = method_colors) +
  geom_point(aes(x = actuals[i_upper], y = 0, shape = "Actual value"), color = "black", size = 3) +
  scale_shape_manual(values = c("Actual value" = 17)) +
  scale_x_continuous(labels = function(x) paste0(formatC(x / 1e3, format = "g"), "k")) +
  guides(fill = "none") +
  labs(
    title = "Predictive densities of upper time series",
    x = "",
    y = ""
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "top",
    legend.title = element_blank()#,
    # axis.title = element_blank(),       
    # axis.text = element_blank(),       
    # axis.ticks = element_blank()
  )
```
Note that all reconciliation methods improve the forecast: the mode of the densities is closer to the actual value. However, the t-Reconciliation method returns a wider density, which is more likely to contain the actual value. This is because t-Reconciliation accounts for the uncertainty in the covariance matrix of the forecast errors, which leads to a more realistic representation of the forecast uncertainty.

# Comparison of the covariance matrix estimates

In this section we showcase the main strength of the t-Reconciliation method: an estimate of the covariance which accounts for the uncertainty in the estimation. To illustrate this point, we compare the covariance matrix estimates obtained with  t-Reconciliation and with the standard Gaussian reconciliation method (MinT).

We focus on the covariance matrix between the upper-level series, "Schweiz" denoted as CH, and the bottom-level time series "Graubünden", denoted as GR, the bottom-level series with the largest average values. Analogous considerations apply also for other series. The code in the vignette is parametric so that a user can manually visualize different comparisons. 

```{r compute correlations, echo=FALSE, eval=TRUE}
# Select which time series to take
i=1     # CH
j=19    # GR

corr_mint <- cov2cor(MinT_cov_mat)
corr_base <- cov2cor(Base_cov_mat)
corr_tRec <- cov2cor(t_rec_results$posterior_Psi/(t_rec_results$posterior_nu - n - 1))
```
```{r,echo=FALSE,eval=TRUE}
# Print table with values for Std and correlation
knitr::kable(data.frame(
  Method = c("Base", "MinT", "t-Rec"),
  Std_CH = sqrt(c(Base_cov_mat[i, i], MinT_cov_mat[i, i], t_rec_results$posterior_Psi[i, i]/(t_rec_results$posterior_nu - n - 1))),
  Std_GR = sqrt(c(Base_cov_mat[j, j], MinT_cov_mat[j, j], t_rec_results$posterior_Psi[j, j]/(t_rec_results$posterior_nu - n - 1))),
  Correlation_CH_GR = c(corr_base[i, j], corr_mint[i, j], corr_tRec[i, j])
), digits=c(0,0,0,2), format.args = list(big.mark = ",",
  scientific = FALSE),caption ="**Table 1**: Variance and correlation estimates for the upper-level series (CH) and the bottom-level series (GR) obtained with the Base, MinT and t-Rec methods.")
```

**Table 1** shows the variance and correlation estimates for the upper-level series (CH) and the bottom-level series (GR) obtained with the Base, MinT and t-Rec methods.Note that the average variance of CH and GR estimated by t-Rec is higher than both the base and the Gaussian reconciled matrix. Moreover, the correlation between CH and GR is also estimated different in t-Rec compared to MinT. 

Most importantly, however, t-Rec returns a distribution for those variance and correlation values. We can plot them by generating samples from the posterior of the covariance. This posterior distribution is an Inverse-Wishart distribution, with parameters returned by `reconc_t()` (if `return_params` is set to `TRUE`) as `posterior_nu` and `posterior_Psi`. We sample from this distribution using the custom function `rinvwishart_base()`, defined below.

```{r generate IW samples}
# generate k samples from an IW(Psi, nu) distribution
rinvwishart_base <- function(k, nu, Psi) {
  p <- nrow(Psi)
  Sigma <- solve(Psi)
  
  all_W <- rWishart(k, df = nu, Sigma = Sigma)
  
  W <- array(NA, dim = c(p, p, k))
  for (i in 1:k) {
    W[,,i] <- solve(all_W[,,i])
  }
  return(W)
}

IW_post_samples <- rinvwishart_base(k=1000, nu = t_rec_results$posterior_nu,
                            Psi = t_rec_results$posterior_Psi)

```

**Figure 3** shows the posterior densities of the standard deviation parameters for CH and GR, and the density of the posterior correlation between CH and GR. The value estimated by MinT, plotted as a vertical dashed line, is not always contained in the support of the densities estimated by t-Rec. For example, for CH, MinT estimates a much lower variance than the one estimated by t-Rec, and this value is in the tail of the density estimated by t-Rec. 

```{r plot densities, echo=FALSE, eval=TRUE, dpi=300, out.width = "100%", fig.align='center', fig.cap="**Figure 3**: Densities of the posterior of the covariance obtained with t-Rec. The vertical line indicates the point estimate obtained with MinT.", fig.dim = c(8, 4)}
# Compute correlation samples for all sample matrices generated above
corr_post <- array(apply(IW_post_samples, FUN = function(M) cov2cor(M), MARGIN = c(3)),
                   dim = dim(IW_post_samples))

name_i <- "CH"
name_j <- "GR"

label_var_i  <- sprintf("sigma[%d%d]", i, i)
label_var_j  <- sprintf("sigma[%d%d]", j, j)
label_cor_ij <- sprintf("rho[%d%d]", i, j)

# display labels with Greek letters + column names
display_labels <- c(
  sprintf("sigma[plain('%s')]", name_i),
  sprintf("rho[plain('%s')*','*plain('%s')]", name_i, name_j),
  sprintf("sigma[plain('%s')]", name_j)
)
names(display_labels) <- c(label_var_i, label_cor_ij, label_var_j)

df <- data.frame(
  x     = c(sqrt(IW_post_samples[i, i, ]), corr_post[i, j, ], sqrt(IW_post_samples[j, j, ])),
  panel = rep(c(label_var_i, label_cor_ij, label_var_j), each = dim(IW_post_samples)[3])
)
df$panel <- factor(df$panel, levels = c(label_var_i, label_cor_ij, label_var_j), ordered = TRUE)

df_mint <- data.frame(
  panel = c(label_var_i, label_cor_ij, label_var_j),
  x     = c(sqrt(MinT_cov_mat[i, i]), corr_mint[i, j], sqrt(MinT_cov_mat[j, j]))
)
df_mint$panel <- factor(df_mint$panel, levels = c(label_var_i, label_cor_ij, label_var_j), ordered = TRUE)

fill_labels <- c(
  setNames(sprintf("sigma[plain('%s')]", name_i), label_var_i),
  setNames(sprintf("rho[plain('%s')*','*plain('%s')]", name_i, name_j), label_cor_ij),
  setNames(sprintf("sigma[plain('%s')]", name_j), label_var_j)
)

cols_fills <- c("#F8766D","#00BA38","#619CFF")

ggplot(df, aes(x = x)) +
  geom_density(aes(fill = panel), alpha = 0.4) +
  geom_vline(data = df_mint,
             aes(xintercept = x, color = "MinT"), linetype = "dashed") +
  scale_fill_manual(
    name   = "",
    values = setNames(cols_fills, c(label_var_i, label_cor_ij, label_var_j)),
    labels = parse(text = fill_labels), #fill_labels,
    guide  = guide_legend(label.hjust = 0)
  ) +
  scale_color_manual(name = "", values = c("MinT" = "black")) +
  facet_wrap(~ panel, nrow = 2, scales = "free",
             labeller = as_labeller(display_labels, default = label_parsed)) +
  labs(x = "", y = "Density") +
  theme_minimal() +
  theme(
    legend.position        = "inside",
    legend.position.inside = c(0.75, 0.2),
    legend.background      = element_rect(fill = "white", color = NA),
    legend.key.width       = unit(1.2, "cm"),
    legend.text            = element_text(size = 8)
  ) +
  guides(
    fill  = guide_legend(override.aes = list(alpha = 0.6)),
    color = guide_legend(override.aes = list(linetype = "dashed"))
  )
```



# References
<div id="refs"></div>

